---
title: | 
  | \vspace{0.25cm} \Large{\textbf{TRIBHUVAN UNIVERSITY}}
  | \vspace{0.20cm} \large{GOKULESHWOR AGRICULTURE AND ANIMAL SCIENCE COLLEGE}
  | \vspace{0.20cm} \large{B. Sc. Ag., Internal Assessment, 2076}
# author: "Deependra Dhakal"
output: 
  bookdown::pdf_document2:
    keep_tex: yes
    toc: false
    include:
      in_header: plb422_biotechnology_biodiversity_internal_assessment_2075-76_header.tex
spacing: singlespacing
fontsize: 11pt
geometry: margin=0.8cm
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, 
                      error=FALSE, message=FALSE)
require(tidyverse)
```

\begingroup
  <!-- \centering -->
  \makebox[4.2cm][l]{\textbf{Subject: Introductory Biotechnology and Biodiversity}}
  \hspace{0.5\textwidth}
  \makebox[1.8cm][r]{\textbf{FM: 10}}
  \par
\endgroup

\begingroup
  <!-- \centering -->
  \makebox[4.2cm][l]{\textbf{Time: 60 Minutes}}
  \hspace{0.5\textwidth}
  \makebox[1.8cm][r]{\textbf{PM: 4}}
  \par
\endgroup

**Level:** $7^{th}$ semester

\textbf{Answer all questions.}

```{r questions, results='asis'}
all_quest <- read_csv("./../../../question_collections/questions_till_date.csv", skip = 1)

# all_quest %>% 
#   select(SN:Question) %>% 
#   filter(Subject == "Introductory biotechnology and biodiversity") %>% 
#   pull(Question) %>% 
#   # Hmisc::latex(file = "", rowname = 1:12, where = 'h')
#   cat(sep = "\\item ")

# # fix this
# cat("\\begin{enumerate}")
# cat(
#   "\\nonumber Essay type (2.5 x 2)", 
#   "\\item Define genetic engineering. Write about its applications in agriculture.", 
#   "\\item What is gene cloning ? Describe about gene cloning vectors. \\\\", 
#   "\\nonumber Short answer type", 
#   "\\item What are restriction enzymes ? Explain their types with suitable examples. (2)", 
#   "\\item What is embryo culture and how is it useful in agricultural crops ? (2)", 
#   "\\item What is species richness ? Calculate Shannon-Wiener index for given insect count data for two different rice fields sampled within a 100 square meter area. (1)"
# )
# cat("\\end{enumerate}")
```

Essay type (2.5 x 2)

1. Define genetic engineering. Write about its applications in agriculture.
1. What is gene cloning ? Describe about gene cloning vectors.

Short answer type

1. What are restriction enzymes ? Explain their types with suitable examples. (2)
1. What is embryo culture and how is it useful in agricultural crops ? (2)
1. What is species richness ? Calculate Shannon-Wiener index for given insect count data for two different rice fields sampled within a 100 square meter area. (1)

```{r questions-data, results='asis'}
tribble(~"Insect", ~"Rice field A", ~"Rice field B", 
        "Gundhi bug", 120, 10, 
        "Hispa", 50, 80,
        "Brown hopper", 80, 85,
        "Grass hopper", 150, 185) %>% 
  knitr::kable(format = "latex", 
               booktabs = TRUE, align = "c") %>% 
  kableExtra::kable_styling(position = "center")
```
