---
title: | 
  | \vspace{0.25cm} \Large{\textbf{TRIBHUVAN UNIVERSITY}}
  | \vspace{0.20cm} \large{GOKULESHWOR AGRICULTURE AND ANIMAL SCIENCE COLLEGE}
  | \vspace{0.20cm} \large{B. Sc. Ag., Internal Assessment, 2076}
# author: "Deependra Dhakal"
output: 
  bookdown::pdf_document2:
    keep_tex: yes
    toc: false
    include:
      in_header: ['gen311_internal_assessment_2075-76_header.tex']
spacing: singlespacing
fontsize: 10pt
geometry: margin=0.8cm
---

```{r setup, include=FALSE}
set.seed(1080)
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, 
                      error=FALSE, message=FALSE)
require(tidyverse)
```

\begingroup
  <!-- \centering -->
  \makebox[4.2cm][l]{\textbf{Subject: Genetics of Populations}}
  \hspace{0.5\textwidth}
  \makebox[1.8cm][r]{\textbf{FM: 10}}
  \par
\endgroup

\begingroup
  <!-- \centering -->
  \makebox[4.2cm][l]{\textbf{Time: 60 Minutes}}
  \hspace{0.5\textwidth}
  \makebox[1.8cm][r]{\textbf{PM: 4}}
  \par
\endgroup

**Level:** $5^{th}$ semester

\textbf{Answer all questions.}

```{r questions, results='asis'}
all_quest <- read_csv("./../../question_collections/questions_till_date.csv", skip = 1)

# cat("\\begin{enumerate} ")
# cat("\\item ")
# all_quest %>%
#   select(SN:Question) %>%
#   filter(Subject == "Introductory genetics") %>%
#   filter(Question_type == "Essay") %>% 
#   pull(Question) %>%
#   sample(1) %>% 
#   # Hmisc::latex(file = "", rowname = 1:12, where = 'h')
#   cat(sep = "\\item ")
# 
# all_quest %>%
#   select(SN:Question) %>%
#   filter(Subject == "Introductory genetics") %>%
#   filter(Question_type == "Short") %>% 
#   pull(Question) %>%
#   sample(4) %>% 
#   # Hmisc::latex(file = "", rowname = 1:12, where = 'h')
#   cat(sep = "\\item ")
# 
# cat("\\end{enumerate}")

```

Essay type (2 x 2 = 4)

1. Write the significance of GxE interaction? Explain types of GxE interaction with diagrams.
2. Explain the model that best explains the structure of chromosome.

Short answer type (1.5 x 4 = 6)

1. What are restriction enzymes? Enlist any 5 restriction enzymes with their cleavage sites. Why is Type-II restriction enzyme family is more useful than others?
2. Explain "one gene one polypeptide hypothesis" with suitable examples if necessary.
3. Explain Johannsens pureline theory. What is meant by 'variegation' in biological tissues?
4. What is recombinant DNA technology? Provide some examples.

\vspace{2cm}

\begingroup
  <!-- \centering -->
  \makebox[4.2cm][l]{\textbf{Subject: Genetics of Populations}}
  \hspace{0.5\textwidth}
  \makebox[1.8cm][r]{\textbf{FM: 10}}
  \par
\endgroup

\begingroup
  <!-- \centering -->
  \makebox[4.2cm][l]{\textbf{Time: 60 Minutes}}
  \hspace{0.5\textwidth}
  \makebox[1.8cm][r]{\textbf{PM: 4}}
  \par
\endgroup

**Level:** $5^{th}$ semester

\textbf{Answer all questions.}

Essay type (2 x 2 = 4)

1. Write the significance of GxE interaction? Explain types of GxE interaction with diagrams.
2. Explain the model that best explains the structure of chromosome.

Short answer type (1.5 x 4 = 6)

1. What are restriction enzymes? Enlist any 5 restriction enzymes with their cleavage sites. Why is Type-II restriction enzyme family is more useful than others?
2. Explain "one gene one polypeptide hypothesis" with suitable examples if necessary.
3. Explain Johannsens pureline theory. What is meant by 'variegation' in biological tissues?
4. What is recombinant DNA technology? Provide some examples.