---
title: Biometrical techniques in plant breeding
author: Deependra Dhakal
institute: |
  | Agriculture and Forestry University
  | \textit{ddhakal.rookie@gmail.com}
  | \url{https://rookie.rbind.io}
output: 
  binb::metropolis:
    includes:
      in_header: beamer_header_template.tex
classoption: "aspectratio=169" # "compressed"
fontsize: 11pt
bibliography: [../bibliographies.bib]
---

```{r,setup, include=FALSE}
library(knitr)
require(tidyverse)
set.seed(453)
# invalidate cache when the package version changes
knitr::opts_chunk$set(tidy = FALSE, echo = FALSE, 
                  message = FALSE, warning = FALSE,
                  out.width = "45%", cache = TRUE)
options(knitr.table.format = "latex")
options(knitr.kable.NA = "", digits = 2)
options(kableExtra.latex.load_packages = FALSE)
```

# Introduction

##

- Ultimate goal of plant breeding is to generate productive cultivars improved for one or more traits.
- Breeding process comprises:
  - Selection of parents possessing desired attributes;
  - Hybridization
  - Progeny selection via series of selection stages

```{r categorization-selection-stages, fig.pos="H", fig.cap="Stages of selection", fig.align="center"}
DiagrammeR::mermaid(
  "
  graph TB
  
  x11[Early generation selection<br>Small amount of large number of lines<br>Practice positive selection] --> x21[Later generation selection<br>Small number of lines<br>Practice negative selection]
  x21 --> x31[One or few released varieties after evaluation in MET]
  
  "
)
```


# Assessment of variability

```{r maize-inbred-exp}

# Source: Page 691; Chapter 19: A Simple Genetic Model for Quantitative Traits; An Introduction to Genetic Analysis, 2012.

maize_inbredexp <- readxl::read_xlsx("./data/maize_inbredline_experiment.xlsx")

maize_inbredexp %>% 
  knitr::kable(caption = "Simulated data for Days to Polllen shed for 10 inbred lines of maize grown in two experiments.")
```

\begin{alert}
Experiment (Table \ref{tab:maize-inbred-exp}) shows simulated data for 10 inbred strains of maize that were grown in three different environments and scored for the number of days between planting and the time that the plants first shed pollen. The overall mean is 70 days. Let's consider line A when grown in environment 1. The mean for all lines in environment 1 is 68, or 2 less than the overall mean, so $e$ for environment 1 is -2. The mean line A over all 3 environments is 64, or 6 less than the overall mean, so g for line A is -6. Putting these two values together, we decompose the phenotype of line A when grown in environment 1 as:

$$
62 = 70 + (-6) + (-2)
$$

\end{alert}

<!-- We could do the same calculations for the other nine inbred lines, and then we'd have a complete description of all the phenotypes in each environment in terms of the extent to which their deviation from the overall mean is due to genetic and environmental factors. -->

## Variance

- We can use the simple model $x = g + e$ to think further about the variance of quantitative traits.
- Variance is a way to measure how much individuals deviate from the population mean. Under this mode, the trait variance can be partitioned into genetic and the environmental variances.

$$
V_x = V_g + V_e
$$

(Continue from the textbook.)

<!-- ## Pooled variance -->

<!-- - Pooled variance (also known as combined, composite, or overall variance; square root known as pooled standard deviation) is a method for estimating variance of several different populations when the mean of each population may be different, but one may assume that the variance of each population is the same. -->
<!-- - Under the assumption of equal population variances, the pooled sample variance provides a higher precision estimate of variance than the individual sample variances.  -->
<!-- - This higher precision can lead to increased statistical power when used in statistical tests that compare the populations, such as the t-test. -->

<!-- <!-- The pooled variance is an estimate of the fixed common variance ${\displaystyle \sigma ^{2}}$ underlying various populations that have different means. --> -->

<!-- ## Computation -->

<!-- If the populations are indexed ${\displaystyle i=1,\ldots ,k}$ then the pooled variance ${\displaystyle s_{p}^{2}}$ can be computed by the weighted average. -->

<!-- $$ -->
<!-- {\displaystyle s_{p}^{2}={\frac {\sum _{i=1}^{k}(n_{i}-1)s_{i}^{2}}{\sum _{i=1}^{k}(n_{i}-1)}}={\frac {(n_{1}-1)s_{1}^{2}+(n_{2}-1)s_{2}^{2}+\cdots +(n_{k}-1)s_{k}^{2}}{n_{1}+n_{2}+\cdots +n_{k}-k}},} -->
<!-- $$ -->

<!-- where ${\displaystyle n_{i}}$ is the sample size of population ${\displaystyle i}$ and the sample variances are -->

<!-- $$ -->
<!-- {\displaystyle {\frac {1}{n_{i}-1}}\sum _{j=1}^{n_{i}}\left(y_{j}-{\overline {y_{i}}}\right)^{2}} -->
<!-- $$ -->

<!-- Use of ${\displaystyle (n_{i}-1)}$ weighting factors instead of ${\displaystyle n_{i}}$ comes from Bessel's correction. -->

<!-- ## Variants -->

<!-- The unbiased least squares estimate of ${\displaystyle \sigma ^{2},}$ -->

<!-- $$ -->
<!-- {\displaystyle s_{p}^{2}={\frac {\sum _{i=1}^{k}(n_{i}-1)s_{i}^{2}}{\sum _{i=1}^{k}(n_{i}-1)}},} -->
<!-- $$ -->
<!-- and the biased maximum likelihood estimate -->

<!-- $${\displaystyle s_{p}^{2}={\frac {\sum _{i=1}^{k}(n_{i}-1)s_{i}^{2}}{\sum _{i=1}^{k}n_{i}}},}$$ -->

<!-- are used in different contexts. The former can give an unbiased ${\displaystyle s_{p}^{2}}$ to estimate ${\displaystyle \sigma ^{2}}$ when the two groups share an equal population variance. The latter one can give a more efficient ${\displaystyle s_{p}^{2}}$ to estimate ${\displaystyle \sigma ^{2}}$ biasedly. Note that the quantities ${\displaystyle s_{i}^{2}}$ in the right hand sides of both equations are the unbiased estimates. -->

<!-- For further details, -->

<!-- https://en.wikipedia.org/wiki/Pooled_variance -->

## Aids to selection 

## Choice of parents

- Appropriate parental combinations for crossing require genotypes of different selection origin and crossing series be screened.
- Evaluation of potential parents require several years [@brown1998identifying].
- Selection of superior genotypes in early generation is highly ineffective for several crops, especially those clonally propagated.
- Each breeding cycle is preceded by (for efficient selection):
  - Cross prediction trials
  - Progeny tests for family selection
- Only a few of all potential hybrid combinations are actually made and evaluated in progeny trials.
- Typically the performance of a new or untested cross combination is predicted by calculating mid-parent values (MPV) based on observed records of potential parents.
- Improved estimates of parental means for a trait are often obtained with some form of an additive linear model.
  - Recently, to address the typical incomplete and unbalanced nature of cross appraisal database, mixed model alternatives to analysis are being used.

## Crossing techniques

## Genotype-by-environment interactions
	
# PART III
	
## Regression Analysis of Stability Finlay and Wilkinsons Model
	
## Analysis of Stability Parameters
	
# PART IV
	
## Griffings Model of Diallel Analysis
	
## Gardner and Eberharts Model of Diallel Analysis
	
## Edwards and Emaras Model of Analysis of Population
	
## Kearseys Model of Biparental Progeny Analysis
	
## North Carolina Designs of Biparental Mating
	
## Factorial Analysis of Generation Means
	
## Detection of Epistasis and Variance Components
	
# PART V

## Line x Tester Analysis Models for Initial Screening
	
## Partial Diallel Analysis Circulent Design
	
## Fractional Diallel Analysis Triangular Designs
	
## Tests of Additivity Epistasis
	
## Genetic parameters
	
## Statistical Genetical Parameters in Mutation Experiments
