---
title: Breeding methods in self pollinated crops
author: Deependra Dhakal
date: Academic year 2019-2020 
institute: |
  | Gokuleshwor Agriculture and Animal Science College
  | Tribhuwan University
  | \textit{ddhakal.rookie@gmail.com}
  | \url{https://rookie.rbind.io}
output: 
  binb::metropolis:
    includes:
      in_header: beamer_header_template.tex
classoption: "aspectratio=169" # "compressed"
fontsize: 11pt
bibliography: [./../../bibliographies.bib]
---

```{r,setup, include=FALSE}
library(knitr)
require(tidyverse)
set.seed(453)
# invalidate cache when the package version changes
knitr::opts_chunk$set(tidy = FALSE, echo = FALSE, 
                  message = FALSE, warning = FALSE,
                  out.width = "45%", cache = TRUE)
options(knitr.table.format = "latex")
options(knitr.kable.NA = "", digits = 2)
options(kableExtra.latex.load_packages = FALSE)
```

# Introduction

##

# Breeding for Rice improvement 

##

- Lecture 14, Cornell University lecture series on Genetic Improvement of Crop Plants -- https://www.youtube.com/watch?v=p-jsjIpUFTE
- Phylogentic relationship between different rice subspecies and groups -- Molecular evaluation of genetic diversity and association studies in Rice (Oryza sativa L.) [@vanniarajan2012molecular]
- Genetic structure and diversity in Oryza sativa L. [@garris2005genetic]

## Types of crosses

- Single crossess
  - P1 x P2
- Backcrosses
  - F1 or F2 are crossed to one of the original parents
- Top cross (Three way cross)
  - F1 x P3
- Double cross
  - F1 x F1

\footnote[frame]{\scriptsize Practical reason for employing more than two parents i.e. using three way or double cross is in order to obtain a larger amount of seed. For e.g., if P1 and P3 do not combine readily or yield only few amount of seed, a mediator line P2 can first be crossed with the P1 to fetch adequate amount of seeds, which can then be planted for hybridization with P3.}

## Characteristics of rice breeding

- Systematic and clear selection criteria
- High-volume crossing (more so if you don't know the parents)
- Effective screening of materials
- Early identification of superior lines (progenies start to accumulate homozygosity very soon)
- Multi location testing and nursery sites

## Inbred breeding methods

- Pedigree (most widely used)
- Bulk breeding
- Single seed descent (SSD)
- Anther culture
- Backcross breeding

## Selection criteria

```{r}
tribble(
  ~Generation, ~`Plant type`, ~`Grain shape`, ~`Maturity`, ~`Height`, ~`Disease resistance`, ~`Yield`, ~`Milling quality`, ~`Cooking quality`,
  "F1", NA, NA, NA, NA, NA, NA, NA, NA,
  "F2", "X", "X", "X", NA, NA, NA, NA, NA,
  "F3", "X", "X", "X", NA, NA, NA, NA, NA,
  "F4", "X", "X", "X", NA, "X", NA, NA, NA,
  "F5", "X", "X", "X", "X", "X", "X", "X", "X",
  "F6", "X", "X", "X", "X", "X", "X", "X", "X",
  "F7", "X", "X", "X", "X", "X", "X", "X", "X"
)
```

## Other breeding methods

- Hybrid breeding
- Transgenic
- Mutation breeding
- Pureline selection
- Molecular breeding
- Doubled haploids

## Hybrid rice breeding

- Refer to 
  - Heterosis and Hybrid Rice Breeding, @virmani1994heterosis 
  - Two-line hybrid rice breeding manual, @virmani2003two

