---
title: Quality seed
author: Deependra Dhakal
date: Academic year 2019-2020 
institute: |
  | Gokuleshwor Agriculture and Animal Science College
  | Tribhuwan University
  | \textit{ddhakal.rookie@gmail.com}
  | \url{https://rookie.rbind.io}
output: 
  binb::metropolis:
    includes:
      in_header: beamer_header_template.tex
classoption: "aspectratio=169" # "compressed"
fontsize: 11pt
bibliography: [./../../bibliographies.bib]
---

```{r,setup, include=FALSE}
library(knitr)
require(tidyverse)
set.seed(453)
# invalidate cache when the package version changes
knitr::opts_chunk$set(tidy = FALSE, echo = FALSE, 
                  message = FALSE, warning = FALSE,
                  out.width = "45%", cache = TRUE)
options(knitr.table.format = "latex")
options(knitr.kable.NA = "", digits = 2)
options(kableExtra.latex.load_packages = FALSE)
```

# Introduction

## What is seed ?

- Defined as a fertilized ovule consisting of intact embryo, stored food and seed coat which is viable and has got the capacity to germinate.
- In crop production, seed material includes tubers, seedling, bulbs, rhizomes, roots and stem cuttings, all kinds of grafts and vegetatively propagated new plants with exact characters of parents.

##

- In some crops, seeds may develop without fertilization or there is no fusion of gametes such process is called apomixis. An apomictic embryo develops simply by division of a diploid cell of the ovule.

## Difference between seed and grain

```{r seed-vs-grain}
seed_vs_grain <- read_csv("../../data/seed_vs_grain.csv") %>% 
  select(-1)

seed_vs_grain %>% 
  knitr::kable(caption = "Difference between grain and seed", booktabs = TRUE) %>% 
  kableExtra::kable_styling(font_size = 6, position = "center", 
                            latex_options = "striped") %>% 
  kableExtra::column_spec(1:2, width = c("25em", "25em"))
```

##

# Bibliography

## See also

## References
