---
title: "Seed quality"
author: "Deependra Dhakal"
date: "2019/04/15 (updated: `r Sys.Date()`)"
output:
  xaringan::moon_reader:
    lib_dir: libs
    self_contained: false # if true, fonts will be stored locally
    seal: false # show a title slide with YAML information
    css: ['xaringan-themer.css', 'assets/slides-style.css', 'assets/custom-remark.css', 'assets/custom-fonts.css']
    chakra: libs/remark-latest.min.js # for offline, use chakra and xaringan::summon_remark()
    nature:
      ratio: "16:9"
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      navigation:
        scroll: false # disable slide transitions by scrolling
      slideNumberFormat: |
        <div class="progress-bar-container">
          <div class="progress-bar" style="width: calc(%current% / %total% * 100%);">
          </div>
        </div>
bibliography: ['../bibliographies.bib']
---

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
# color inspiration: pantone 2020 colors
# https://store.pantone.com/uk/en/color-of-the-year-2020-palette-exploration
style_duo_accent(
  primary_color = "#0F4C81",   # pantone classic blue
  secondary_color = "#DBCCBE", # pantone pink tint
  black_color = "#7A8387",     # pantone monument
  text_color = "#2A2A35",      # pantone night sky
  header_font_google = google_font("Arvo"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Source Code Pro"),
  title_slide_text_color = "#DBCCBE",
  base_font_size = "24px",
  text_font_size = "0.8rem",
  link_color = "#B6CADA"
)
```

class: inverse, middle, center
background-color: blue
background-size: 60%, auto

<style type="text/css">
.remark-slide-content {
    font-size: 16px;
    padding: 1em 4em 1em 4em;
}
</style>

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
options(knitr.kable.NA = "", digits = 3)
knitr::opts_chunk$set(tidy = FALSE, echo = FALSE, warning = FALSE, message = FALSE)
require(tidyverse)

# pagedown::chrome_print("./14-breeding_methods_in_cross_pollinated_crops_html.html")
# better pdf rendering with manually printing of chromium
```

# Seed systems in Nepal

---

Broadly, two types of seed system are recognized in Nepal

1. Informal Seed System
  - Characterized by farmers producing and preserving their own seeds for subsequent planting.
  - Farmers exchange small amount of seeds with other farmers as gift, and for both monetary and non monetary value.
  - Most traditional and local landraces are product of such selection and maintenance process, these land races are important genetic resources for modern plant breeding. 
  - Seed production process is usually managed as a part of crop production and seed distribution as per farmers' choice and demand among neighbors. 
  - Use of retained seeds from their own production, farmers to farmers' seed exchange, informal purchase, gifts etc.

2. Formal seed system
  - Characterized by a vertically organized production and distribution of tested and released/registered varieties by public and private organizations using agreed quality control mechanism.
  - Comprises different phases of seed cycle: Breeder, Foundation, Certified and Improved seeds. Formal system: It comprises public or private sector organizations that are linked to the national seed system. NARC, DOA, NSC, Seed companies, NGOs, DISSPRO and CBSP groups etc are under this category.

---
class: inverse, center, middle
background-color: green

# Seed versus Grain

---

```{r}
readr::read_csv("./data/seed_vs_grain.csv") %>% 
  knitr::kable(booktabs = TRUE)
```

---

## Seed policy, 1999

- Ownership of traditional varieties and conservation of Agrobiodiversity
- Technical support in seed production and supply by organizing farmers groups in remote areas to enhance access of quality seeds to small farmers
- Subsidy on source seed production and transportation of seeds to remote hills and mountain regions; seed supply will be expanded in coordination with private producers and traders including by giving some support.
- Involvement of private sectors in source seed production and seed supply in remote areas and seed quality assurance services
- Research on Biotechnology and regulation of GMO; Research and studies on biotechnology, genetic engineering, GMO, transgenic plant and tissues culture will proceed in Nepal also but biotechnology regulations will be prepared and implemented for the sake of the general public.
- Seed supply: There will be a buffer stock of seed at the national level, which may be utilized under conditions of natural calamity.
- Institutional strengthening: This will be fulfilled by the strengthening of the NSB Secretariat, and establishment of non-government laboratories, besides including institutional management in the contracting system whilst the semi-government agency involved in seed will be commercialized.

---

## National Seed Vision (2013-2025)

- Self-sufficiency, import substitution and export promotion
- Conservation of indigenous varieties and traditional knowledge for crop improvement (sustainable use)
- Formulate and amend policy and act for ensuring farmers' rights on seeds
- Develop policy for breeders rights to provide incentives and investment in competent seed industry development
- Provision of contract farming act including seed insurance and mechanization for seed production
- Harmonization of seed policy with other Nepalese Policies, laws and also with neighbouring countries
- Focus on hybrid variety development to provide adequate options and access of HYVs to farmers by encouraging both public and private sectors
- Simplify variety approval, release and registration of indigenous and farmers varieties
- Decentralize source seed production for enhancing access to quality seeds and strengthening capacity of private sector and communities in seed production and supply.
- Public-private partnership (Public sector in breeding and private sector in marketing)

---
class: inverse, middle, center

# Classes of seed

---

1. Breeder seed
  - Genetic purity of the variety: 99.9%
  - Golden yellow tag (as of new seed certification guidelines, it is ?brown) with breeder's signature.
2. Foundation seed
  - Not available for sales/purchase at farmer's level
  - SQCC or specialized organization having granted permit from the SQCC performs the field inspection and certification
  - White colored tag with black letters
2. Certified seed
  - In self pollinated crops two generations may be grown, however only one generation is allowed in cross pollinated species
  - CS-I and CS-II (Blue letters in white tag and blue border, and green letters in white tag with green borders, respectively)
3. Improved seed
  - In cross pollinated and highly CP species, direct production of improved seed from foundation seed is also practiced (Maize, vegetable, pigeonpea).
  - Production takes place in farmer's field or by seed producing organizations.
  - Yellow colored tag

---

.pull-left[
- Normally, genetic purity of FS is 99.5 and that of CS is 99%.
- For longer storage in cereals, vegetables, legumes and oilseeds, and orthodox seeds, optimum moisture content: 12% or less.
- In recalcitrant seeds, moisture content is best kept close to 20%.
- In the moisture regime between 5-14%, for every drop in 1% storage life of the seed is doubled (Jems F Herington)
- Forbidden diseases of crops:
  - Cauliflower: Black rot, Alternaria leaf spot
]

.pull-right[
```{r number-of-seed-per-ten-g}
tibble::tribble(~"Crop", ~"Number of seeds", 
        "Carrot", 8280, 
        "Chinese cabbage", 6480,
        "Turnip", 5400, 
        "Tomato", 4140,
        "Onion", 3420, 
        "Pea", 3263,
        "Cabbage, cauliflower", 3240,
        "Brinjal", 2340,
        "Capsicum", 1620,
        "Radish", 1440,
        "Chukandar", 576,
        "Muskmelon", 470,
        "Cucumber", 400,
        "Okra", 180,
        "Watermelon", 75,
        "Luffa gourd", 40
        ) %>% 
  knitr::kable(booktabs = TRUE, caption = "Number of seeds per 10 gram of vegetable crops") %>% 
  kableExtra::kable_styling(latex_options = "striped", font_size = 10)
```
]

---

### Seed certification system

.pull-left[
- Seed act was first enacted in 2045 BS. Since then it's first amendment came into being in 2064.
- Organized seed production program intitiated in Nepal in 1973 (Wheat)
- 90% of the seeds are supplied through informal sector (SQCC, 2011).
- Seed replacement rate of cereals: 9.75% (SQCC, 2011)
- SRR for vegetable crops was as high as 66% in 2009 (NSV)
- Establishment of agriculture input corporation: 1974
- In 1960, High yielding variety of wheat Lerma-52 was released.
- In 1962, Seed testing laboratory was established under Agronomy Division, Khumaltar.
- In 1964, Central Seed Lab of Nepal was designated membership with International Seed Testing Association (ISTA).
- In 1966, Seed testing laboratory moved to Agri-Botany Division.
- In 1966, Rice variety CH-45 was released.
- In 1974, AIC established under Corporation Act, 1965.
- In 1980, Seed production and input storated project (SPISP, USAID) launched
]

.pull-right[
- In 1982, Seed technology and improvement programme
- In 1984, Central seed science and technology division established
- In 1988, Seed act enacted
- In 1997, National seed policy approved
- In 1998, Seed production guidelines published
- In 2001, Seed quality control center established
- In 2002, National seed company limited established
- In 2010, Seed science and technology division revived as central disciplinary division of seed in NARC
- In 2013, Promulgation of seed regulation, 2013
- In 2004, First hybrid of maize -- Gaurav, was officially released.
- In 2010, First hybrid of tomato -- Srijana, was registered
- Upto 2010, 231 hybrids were registered for cultivation in Nepal.
]

---

**Two certification/authentication system are practiced in Nepal**

1. Seed certification: Official body responsible for certifying seed in the SQCC. Three classes of seeds are identified: Foundation, Certified-I, Certified-II and Improved under this system.
2. Truthful labelling

- Following activities are carried out in various stages of a crop are crucial for inspection by a seed inspector:
  - Standing crop: By licenced inspector, for field inspection check for crop's source and seed class, purity, isolation distance, seed affecting diseases, weed and type of varieties.
  - At harvest: Threshing, processing, transporation and storage.
  - After processing: To verify that minimun quality meets, insect damage (In legume: and maize: 1% and other crops: 0.5%), seed moisture, weed and off type seeds, germination percentage
  - Seed is sampled and taken to lab
  - Lab should return results within 30 days of receiving a sample.
  - If suspected, sample may be subjected to grow out test by the tester.
  - If farmer isn't convinced of the results of the seed testing, s/he may file for recertification.

- Tagging storage container with certification tag, with signature of certifying inspector.
- If the seed is to be stored for more than 1 season, the storage sample must be re-certified for quality standards.
- There are some steps to be followed in prolonging the validity of certification. 

---

1. Minimum standard certification:
  - Location and land requirement fulfilment certification
    - In terai, a minimum of 1 hectares is required for cereal and cash crops.
    - In terai, for vegetables, a minimum of 0.25 hectares is required.
    - In hills, for cereal, a minimum of 0.25 hectare is the necessary criterion.
    - In hills, for vegetables, a minimum of 0.1 ha land should be cultivated with seed under the question.
    - Maximum distance between plots of same certification lot should not exceed 50 m.
2. Location and locality of specification
3. Source of seed verification
4. Timing of inspector assignment

---

### Certification standards

```{r certification-standards-field}
readxl::read_xlsx("./data/seed_certification_requirements.xlsx", sheet = "field_standards") %>% 
  rename_all(function(x)str_replace_all(x, "_", " ")) %>% 
  knitr::kable(caption = "Field standards of standing crop for certification", booktabs = TRUE, longtable = TRUE) %>% 
  kableExtra::kable_styling(latex_options = "striped", font_size = 8) %>% 
  kableExtra::column_spec(column = 1:9, width = c("8em", rep("6em", times = 8))) %>% 
  kableExtra::landscape()
```

---

```{r certification-standards-seed}
readxl::read_xlsx("./data/seed_certification_requirements.xlsx", sheet = "seed_standards") %>%
  select(1:9) %>%
  rename_all(function(x)str_replace_all(x, "_", " ")) %>% 
  knitr::kable(caption = "Minimum seed standards for certification", booktabs = TRUE) %>% 
  kableExtra::column_spec(column = 1:9, width = c("4em", rep("5em", times = 8)))
# kableExtra::landscape(margin = c("1cm", "1cm", "1cm", "1cm"))
readxl::read_xlsx("./data/seed_certification_requirements.xlsx", sheet = "seed_standards") %>% 
  select(1,10:15) %>% 
  rename_all(function(x)str_replace_all(x, "_", " ")) %>% 
  knitr::kable(caption = "Minimum seed standards for certification", booktabs = TRUE) %>% 
  kableExtra::column_spec(column = 1:7, width = c("4em", rep("6em", times = 6)))
```

