---
title: "Biodiversity indexing"
author:   
  - Deependra Dhakal 
institute:   
  - GAASC, Baitadi
  - Tribhuwan University 
date: Academic year 2019-2020 
output:   
  beamer_presentation:  
    incremental: false  
    theme: "Frankfurt"  
    colortheme: "beaver"
    toc: true   
    slide_level: 3
    keep_tex: true
    includes:
      in_header: 08-biodiversity_indexing_beamer_header.tex
classoption: "aspectratio=169"
header-includes: 
- \AtBeginSubsection{}
bibliography: [./../bibliographies.bib]
---

```{r setup, include=FALSE}
library(knitr)
require(tidyverse)
set.seed(453)
# invalidate cache when the package version changes
knitr::opts_chunk$set(tidy = FALSE, echo = FALSE, 
                  message = FALSE, warning = FALSE,
                  out.width = "45%")
options(knitr.table.format = "latex")
options(knitr.kable.NA = "", digits = 2)
options(kableExtra.latex.load_packages = FALSE)
```

# Indices

## 

$$
{}^{q}\!D=\left({\sum _{i=1}^{R}p_{i}^{q}}\right)^{1/(1-q)}
$$

and the term inside the parentheses is called the basic sum. Some popular diversity indices correspond to the basic sum as calculated with different values of q.

## Richness

# Numerical problems

## Problem

1. Calculate Shannon Weaner Index, Simpson Index and Evenness from following two areas and interpret your result. Also discuss species richness and evenness from following information.

```{r}
tribble(~"Insect", ~"Area 1", ~"Area 2", 
        "Wasp", 112, 424,
        "Butterfly", 88, 76, 
        "Grasshopper", 143, 54, 
        "Bettle", 112, 60, 
        "Bee", 100, 40, 
        "Hoverfly", 145, 46) %>% 
  t() %>% 
  magrittr::set_colnames(.[1,]) %>% 
  as_tibble() %>% 
  slice(-1)
```

## Solution

Here, 

We have Shannon-Weiner Index,

$$ 
H = \sum_{i = 1}^S{p_i.\ln p_i}
$$

Where, $S$ is the species richness (number of distinct species types), $p_i$ is the proportional abundance, and $\ln p_i$ the natural log of the proportional abundance.

```{r}
tribble(~"Insect", ~"Area 1", ~"Area 2", 
        "Wasp", 112, 424,
        "Butterfly", 88, 76, 
        "Grasshopper", 143, 54, 
        "Bettle", 112, 60, 
        "Bee", 100, 40, 
        "Hoverfly", 145, 46) %>% 
  mutate_at(c("Area 1", "Area 2"), list(pi = ~./(sum(.)), lnpi = ~log(./(sum(.))))) %>% 
  summarise(A1_d_shannon_weiner = -sum(`Area 1_lnpi` * `Area 1_pi`), 
            A1_d_simpson = -sum(`Area 1_pi`^2), 
            A1_d_difference_simpson = 1-A1_d_simpson,
            A1_d_richness = n(),
            A1_d_evenness = A1_d_shannon_weiner/log(A1_d_richness),
            A2_d_shannon_weiner = -sum(`Area 2_lnpi` * `Area 2_pi`), 
            A2_d_simpson = -sum(`Area 2_pi`^2), 
            A2_d_difference_simpson = 1-A2_d_simpson, 
            A2_d_richness = n(),
            A2_d_evenness = A2_d_shannon_weiner/log(A2_d_richness)
            ) %>% 
  pivot_longer(cols = A1_d_shannon_weiner:A2_d_evenness, 
               names_to = c("Area", "Index"), 
               names_pattern = "A(\\d)_d_(.*)"
               ) %>% 
  pivot_wider(names_from = Area, names_prefix = "Area ")
```


