---
title: "Practice Set - I"
author: |
  | Deependra Dhakal
# date: "9/18/2020"
output: 
  bookdown::pdf_document2:
    keep_tex: yes
    toc: false
documentclass: exam
classoption: "answers,addpoints,12pt"
geometry: "top=1.5cm,bottom=1.5cm,left=1.5cm,right=1.5cm"
header-includes:
  - \newcommand{\bquestions}{\begin{questions}}
  - \newcommand{\equestions}{\end{questions}}
  # - \newcommand{\bsolutionex}{\begin{solution}}
  # - \newcommand{\esolutionex}{\end{solution}}
  - \newcommand{\bparts}{\begin{parts}}
  - \newcommand{\eparts}{\end{parts}}
  - \newcommand{\stpart}{\part} % this is absolutely necessary to make new part command
  - \newcommand{\bsubparts}{\begin{subparts}}
  - \newcommand{\esubparts}{\end{subparts}}
  - \newcommand{\stsubpart}{\subpart}
  - \usepackage{float} % this package is essential for solution with code
  - \usepackage{eqnarray, amsmath}
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{wrapfig}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{tabu}
  - \usepackage{threeparttable}
  - \usepackage{threeparttablex}
  - \usepackage[normalem]{ulem}
  - \usepackage{makecell}
  - \usepackage{xcolor}
  - \usepackage{amsmath}
  - \usepackage{caption}

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, include = TRUE, message = FALSE, warning = FALSE)
options(knitr.kable.NA = '')
require(tidyverse)
```

\textbf{Numerical Problems}

\bquestions

\question \label{quest:first}

How often in a family of 3 children where both parents are heterozygous for non-red hair (Rr) and have wavy hair (h1h2) will the children be:

i. A red wavy-haired girl, a non-red curly-haired boy and a red straight-haired girl
ii. Two red wavy-haired boy and one non-red straight-haired girl

<!-- \bsolutionex -->

(Solution to Question \ref{quest:first}) 
$\longrightarrow$

```{r}
choose(6, 2)
combn(letters[1:6], 2)

```


So, 


<!-- \esolutionex -->

\equestions
