---
title: Biometrical techniques in plant breeding
author: Deependra Dhakal
date: Academic year 2019-2020 
institute: |
  | Gokuleshwor Agriculture and Animal Science College
  | Tribhuwan University
  | \textit{ddhakal.rookie@gmail.com}
  | \url{https://rookie.rbind.io}
output: 
  binb::metropolis:
    includes:
      in_header: beamer_header_template.tex
classoption: "aspectratio=169" # "compressed"
fontsize: 11pt
bibliography: [./../../bibliographies.bib]
---

```{r,setup, include=FALSE}
library(knitr)
require(tidyverse)
set.seed(453)
# invalidate cache when the package version changes
knitr::opts_chunk$set(tidy = FALSE, echo = FALSE, 
                  message = FALSE, warning = FALSE,
                  out.width = "45%", cache = TRUE)
options(knitr.table.format = "latex")
options(knitr.kable.NA = "", digits = 2)
options(kableExtra.latex.load_packages = FALSE)
```

# Introduction

##

```{r maize, fig.cap="Biometrical techniques in plant breeding", fig.align='center', out.height="400px", out.width="auto"}
# knitr::include_graphics("./images/")
```

---
class: inverse, center, middle

# Assessment of variability

---

# Assessment of variability

## Variance

## Pooled variance

- Pooled variance (also known as combined, composite, or overall variance; square root known as pooled standard deviation) is a method for estimating variance of several different populations when the mean of each population may be different, but one may assume that the variance of each population is the same.

- Under the assumption of equal population variances, the pooled sample variance provides a higher precision estimate of variance than the individual sample variances. 

- This higher precision can lead to increased statistical power when used in statistical tests that compare the populations, such as the t-test.

---

The pooled variance is an estimate of the fixed common variance ${\displaystyle \sigma ^{2}}$ underlying various populations that have different means.

## Computation

If the populations are indexed ${\displaystyle i=1,\ldots ,k}$ then the pooled variance ${\displaystyle s_{p}^{2}}$ can be computed by the weighted average.

$$
{\displaystyle s_{p}^{2}={\frac {\sum _{i=1}^{k}(n_{i}-1)s_{i}^{2}}{\sum _{i=1}^{k}(n_{i}-1)}}={\frac {(n_{1}-1)s_{1}^{2}+(n_{2}-1)s_{2}^{2}+\cdots +(n_{k}-1)s_{k}^{2}}{n_{1}+n_{2}+\cdots +n_{k}-k}},}
$$

where ${\displaystyle n_{i}}$ is the sample size of population ${\displaystyle i}$ and the sample variances are

$$
{\displaystyle {\frac {1}{n_{i}-1}}\sum _{j=1}^{n_{i}}\left(y_{j}-{\overline {y_{i}}}\right)^{2}}
$$

Use of ${\displaystyle (n_{i}-1)}$ weighting factors instead of ${\displaystyle n_{i}}$ comes from Bessel's correction.

## Variants

The unbiased least squares estimate of ${\displaystyle \sigma ^{2},}$

$$
{\displaystyle s_{p}^{2}={\frac {\sum _{i=1}^{k}(n_{i}-1)s_{i}^{2}}{\sum _{i=1}^{k}(n_{i}-1)}},}
$$
and the biased maximum likelihood estimate

$${\displaystyle s_{p}^{2}={\frac {\sum _{i=1}^{k}(n_{i}-1)s_{i}^{2}}{\sum _{i=1}^{k}n_{i}}},}$$

are used in different contexts. The former can give an unbiased ${\displaystyle s_{p}^{2}}$ to estimate ${\displaystyle \sigma ^{2}}$ when the two groups share an equal population variance. The latter one can give a more efficient ${\displaystyle s_{p}^{2}}$ to estimate ${\displaystyle \sigma ^{2}}$ biasedly. Note that the quantities ${\displaystyle s_{i}^{2}}$ in the right hand sides of both equations are the unbiased estimates.

---

## Pooled variance

For further details,

https://en.wikipedia.org/wiki/Pooled_variance

---

# aids to selection 

# choice of parents

# crossing techniques

# genotype-by-environment interactions

---